---
title: "The Fallacy of Placing Confidence in Confidence Intervals"
author:
- name: Richard D. Morey
  affiliation: Cardiff University
- name: Rink Hoekstra
  affiliation: University of Groningen
- name: Jeffrey N. Rouder
  affiliation: University of Missouri
- name: Michael D. Lee
  affiliation: University of California-Irvine
- name: Eric-Jan Wagenmakers
  affiliation: University of Amsterdam
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output: 
  html_document:
    fig_caption: yes 
    dev: CairoSVG
    keep_md: yes
bibliography: bibfile.bib
csl: apa.csl
---

```{r results='hide',echo=FALSE,message=FALSE,warning=FALSE}
source('figureCode/prelims.R', chdir = TRUE)

opts_knit$set(self.contained = FALSE)
```


```{r child="sectionTexts/introduction.Rmd"}
```

## The Folk Theory of Confidence Intervals

```{r child="sectionTexts/folkTheory.Rmd"}
```

## The Theory of Confidence Intervals

```{r child="sectionTexts/neymanTheory.Rmd"}
```

### Early skepticism

```{r child="sectionTexts/earlySkepticism.Rmd"}
```

## Example 1: The Lost Submarine

```{r child="sectionTexts/lostSubIntro.Rmd"}
```

### Five confidence procedures

```{r child="sectionTexts/subProcedures.Rmd"}
```

### Properties of the procedures

```{r child="sectionTexts/subCIproperties.Rmd"}
```

### Evaluating the confidence procedures

```{r child="sectionTexts/subCIevaluating.Rmd"}
```

## Example 2: A confidence interval in the wild

```{r child="sectionTexts/omega2.Rmd"}
```

## Discussion

```{r child="sectionTexts/discussionIntro.Rmd"}
```

### Guidelines for interpreting and reporting intervals

```{r child="sectionTexts/guidelines.Rmd"}
```

### Confidence intervals versus credible intervals

```{r child="sectionTexts/confVcred.Rmd"}
```

### Conclusion

```{r child="sectionTexts/conclusion.Rmd"}
```


# References
